# Project 2: Housing Prices in King Co. Washington, 2014-15

## Exploratory data analysis

Columns: 
id - Unique ID for each home sold
date - Date of the home sale
price - Price of each home sold
bedrooms - Number of bedrooms
bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower
sqft_living - Square footage of the apartments interior living space
sqft_lot - Square footage of the land space
floors - Number of floors
waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not
view - An index from 0 to 4 of how good the view of the property was
condition - An index from 1 to 5 on the condition of the apartment,
grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
sqft_above - The square footage of the interior housing space that is above ground level
sqft_basement - The square footage of the interior housing space that is below ground level
yr_built - The year the house was initially built
yr_renovated - The year of the house’s last renovation
zipcode - What zipcode area the house is in
lat - Lattitude
long - Longitude
sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors
sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors


no missing values, but there is a little housecleaning to do.
houses with >10 bedrooms seem suspect
6 houses list no bedroom but have bathroom (studio? but 3 with 2.5 bath?)
7 with no bed or bath, but >3k sqft and >$1mil

But don't delete until after fitting the model



```{r}
#import data
library(tidyverse)

kc_df<-as_tibble(read.csv("~/MSDS/STAT6021/Project2/kc_house_data.csv"))
```



```{r}
ggplot(kc_df, aes(x=long, y=lat, color=yr_built))+
  geom_hexbin()
```



```{r}
#correlations? pull out the quantitative variables and see
kc_quant<-kc_df %>% select(price, sqft_living, sqft_lot, condition, grade, sqft_above, sqft_basement, sqft_living15, sqft_lot15)
kc_cor<-cor(kc_quant)
kc_cor %>% heatmap()
print(kc_cor)
print(min(kc_cor))
```
Highest correlation (.877) between sqft_above and sqft_living (shouldn't include both in a model bc of collinearity)
```{r}
install.packages("corrplot")
library(corrplot)
```

```{r}
corrplot(kc_cor, method = 'ellipse')#,
#          method = "square",
#          type = "lower",
#          tl.col = "black",
#          tl.cex = 1,
#          col = colorRampPalette(c("brown", "purple"))(100)
# )
```


```{r}

ggplot(kc_df, aes(long, lat, color=2015-yr_built))+
  #geom_density_2d(mapping=aes(color=price), bins=20)+
  geom_point(alpha=.15, color='darkgreen')

#plot(kc_df$long, kc_df$lat, col=ifelse(kc_df$waterfront=='1', "green", ifelse((kc_df$view==4 & kc_df$waterfront=='0'), rgb(red = 1, green = 0, blue = 0, alpha = 1), rgb(red = 0, green = 0, blue = 1, alpha = 0.2))))
```
Green dots are classified 'waterfront', 


```{r}
table(kc_df$waterfront)
```

Properties considered waterfront are very rare in the data (less than 1% of the )


```{r}
table(kc_df$yr_renovated!=0)
```

Almost 1000 houses renovated, but a tiny fraction of the housing stock. 


Do houses on the waterfront have good views?

```{r}
kc_df$waterfront<-factor(kc_df$waterfront)
ggplot(kc_df, aes(x=waterfront, y=view))+
  geom_violin()
```
where are there houses with 4 star views that are not on the waterfront? There are 184 of them. are they clustered?



```{r}
four_star_offwater<-kc_df %>% filter(view==4 & waterfront=='0')
ggplot(four_star_offwater, aes(long, lat))+
  geom_point()


```
Sort of all over the place. Many are likely waterfront, but not so labeled.

What about against the discrete factors

```{r}
ggplot(kc_df, aes(factor(grade), price))+
  geom_boxplot()
```

```{r}
ggplot(kc_df, aes(factor(view), price))+
  geom_boxplot()
```

```{r}
ggplot(kc_df, aes(factor(condition), price))+
  geom_boxplot()
```

```{r}
ggplot(kc_df, aes(factor(zipcode), price))+
  geom_boxplot()
```

One zipcode has the highest median by a lot, but doesn't have the most expensive place.

have houses been getting bigger or smaller over time?

```{r}
ggplot(kc_df, aes(yr_built, sqft_living))+
  geom_point()+
  geom_smooth(method=lm)
```
similarly have the number of bedrooms/bathrooms in new construction gone up?

```{r}
ggplot(kc_df , aes(yr_built, bedrooms))+
  geom_point()+
  geom_smooth(method=lm)
```

```{r}
ggplot(kc_df , aes(yr_built, bathrooms))+
  geom_point()+
  geom_smooth(method=lm)
```

The trend toward more bathrooms definitely seems legit.



```{r}
#convert date to datetime
library(lubridate)
kc_df$date<-as.Date(sub('(\\d{2})(\\d{2})$', '-\\1-\\2', substring(kc_df$date,1,8)))

```

```{r}
class(kc_df$date)
month(kc_df$date[1:5])
```

"Happy families are all alike; every unhappy family is unhappy in its own way."
Using location information, like a bounding box of lat/long, KNN
