---
title: "Question 1"
author: "Tyler Valentine"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages
```{r}
library(tidyverse)
library(leaps)
library(faraway)
```

### Import data

```{r}
data <- read.csv("seattle_housing_data.csv")
```

### A little EDA: distribution of ages of houses in the data
```{r}
ggplot(data, aes(x=(2015-age), y=bathrooms))+
  geom_point(alpha=.2)+
  geom_smooth(method=lm, color="red", se=FALSE)+
  labs(title="Number of bathrooms by year built", x='Year Built')+
  theme(legend.position = "none")
```
Density plot:

```{r}
ggplot(data, aes(x=2015-age))+
  geom_density()+
  labs(title="Density plot of year built", x='Year Built')
```



### Separate training and test data

```{r}
set.seed(206)
sample <- sample.int(nrow(data), floor(.80*nrow(data)), replace = F)
train <- data[sample, ] ##training data
test <- data[-sample, ] ##test data
```

### Linear regression

```{r}
full.model <- lm(age~price+bedrooms+bathrooms+sqft_living+sqft_lot+
                   floors+condition+grade+sqft_above+distance+sqft_living15+
                   sqft_lot15+waterfront+view+basement+renovated+season, data=train)
summary(full.model)
```

### Partial F test to see if we can drop predictors. 

Drop season, sqft_lot_15
```{r}
reduced.model <- lm(age~price+bedrooms+bathrooms+sqft_living+sqft_lot+sqft_above+
                   floors+condition+grade+distance+sqft_living15+ 
                   waterfront+view+basement+renovated, data=train)
full.model <- lm(age~price+bedrooms+bathrooms+sqft_living+sqft_lot+
                   floors+condition+grade+sqft_above+distance+sqft_living15+ sqft_above+
                   sqft_lot15+waterfront+view+basement+renovated+season, data=train)
anova(reduced.model,full.model)
```

Check for multicollinearity:
Drop sqft_above
```{r}
vif(reduced.model)
```

Use VIF to check for multicollinearity in final model. 
```{r}
vif(final.model)
```

```{r}
final.model <- lm(age~price+bedrooms+bathrooms+sqft_living+sqft_lot+
                   floors+condition+grade+distance+sqft_living15+ 
                   waterfront+view+basement+renovated, data=train)
```

### Final model 
```{r}
summary(final.model)
```

### Regression assumptions

Residual plot for final.model

```{r}
yhat<-final.model$fitted.values
res<-final.model$residuals
Data<-data.frame(train, yhat, res)
##residual plot
ggplot(Data, aes(x=yhat,y=res))+
geom_point()+
geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot of Final Model")

```


ACF
```{r}
acf(res, main="ACF of Residuals from Final Model")

```
QQ Plot

```{r}
qqnorm(res)
qqline(res, col="red")
```

## Testing the Model

```{r}
preds<-predict(final.model, newdata = test, interval = 'prediction')

#a function to predict the accuracy using test data
accuracy<-function(predict_vec, threshold){
  sum(abs(predict_vec[,1]-test$age)<=threshold)/dim(predict_vec)[1]}

accuracy(preds, 5)
```

```{r}
accuracy(preds, 10)
```



Accuracy of 22.9% for predictions within 5 years, 43.6% for predictions within 10 years. 

