---
title: "Project 2: Data Cleaning"
author: "Group 9"
date: "08/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages
```{r}
library(tidyverse)
library(corrplot)
library(leaps)
library(faraway)
```

### Data cleaning

Remove unnecessary variables (ID and zip code). 
```{r}
data <- read.csv("~/MSDS/STAT6021/Project2/kc_house_data.csv") 
```

#### Define categorical variables 

Change waterfront to categorical.
```{r}
data$waterfront <- factor(data$waterfront)
levels(data$waterfront) <- c("no","yes") 
```

Change view to categorical. 
```{r}
data$view <- factor(data$view)
levels(data$view) <- c("no", "yes", "yes", "yes", "yes") 
```

Create categorical variable 'renovated':
```{r}
data <- data %>% mutate(renovated = yr_renovated>0)
data$renovated<-factor(data$renovated)
levels(data$renovated)<- c("no","yes") 
#data <- rename(data, renovated = yr_renovated)
```

#### Create new variables 

Change yr_built to age. 
```{r}
data <- mutate(data, age = 2015 - yr_built)

```

Create categorical variable over_budget ($718,000). 
```{r}
data <- data %>% mutate(over_budget = price > 718000)
data$over_budget <- factor(data$over_budget)
levels(data$over_budget) <- c("no","yes") 
```

Change sqft_basement to basement. 
```{r}
data <- data %>% mutate(has_basement = sqft_basement>0)
data$has_basement <- factor(data$has_basement)
levels(data$has_basement) <- c("no","yes") 
```

Change date to season. 
```{r}
data$date <- substr(data$date,start=1,stop=8)
data$year <- substr(data$date,start=1,stop=4)
data$month <- substr(data$date, start=5, stop=6)
data$day <- substr(data$date, start=7, stop=8)
data$season <- ifelse(data$month=="05"|
                       data$month=="06"|
                       data$month=="07"|
                       data$month=="08"|
                       data$month=="09","summer","non-summer")
```

Delete columns we no longer need.
```{r}
data <- subset(data, select = -c(date, year, month, day) )
```

Change order of columns. 
```{r}
data <- data[,c("age",
                "price",
                "over_budget",
        "bedrooms",
        "bathrooms",
        "sqft_living",
        "sqft_lot",
        "floors",
        "condition",
        "grade",
        "sqft_above",
        "lat",
        "long",
        "sqft_living15",
        "sqft_lot15",
        "waterfront",
        "view",
        "has_basement",
        "renovated",
        "season")]
```

### Correlation plot of numerical variables

```{r}
cor <- select(data, !c(over_budget, waterfront, view, has_basement, renovated, season)) %>%
                cor(use = 'complete.obs') %>%
                round(2)

corrplot(cor, type = "lower", order = "FPC", diag = T, method = 'square',
         col = COL2('PuOr'),
         tl.col = "black", tl.srt = 45,
         title = "Correlation matrix of numerical variables", 
         mar=c(0,0,1,0))
```
We will be trying to predict the age of a house based on the other evidence available to us. 
At first glance, we can see that some features show a high negative correlation with age: sqft_above and sqft_living, grade, bathrooms, sqft_living15, floorsm and longitude (east-west position). Latitude and condition show a moderate positive correlation.
Sqft_above and sqft_living are measuring very similar things, and are very highly correlated, so only one will likely be able to stay in the model.

So let's get to it!

Fit a linear model
